Exploring Nonlinear Atmospheric Influences on Temperature Using Polynomial Regression at Vancouver International Airport

### 2024-11-27 paper/paper.qmd
update title and introduction

---
title: "Non-Linear Effects of Pressure, Precipitation, Wind, and Gust Speed on Temperature at Vancouver International Airport"
subtitle: "A Polynomial Regression Approach to Analyzing Atmospheric Influences on Temperature"
author: 
  - Lexun Yu
thanks: "Code and data are available at: [https://github.com/yulexun/ClimateChangeYVR](https://github.com/yulexun/ClimateChangeYVR)."
date: today
date-format: long
abstract: "This study investigates how atmospheric factors influence temperature at Vancouver International Airport using a polynomial regression model. The analysis shows that total precipitation has the strongest effect, significantly reducing mean temperature, while wind speed and atmospheric pressure exhibit non-linear and consistent negative impacts, respectively. By explaining 61% of the variability in temperature, the model demonstrates that precipitation, pressure, gust speed and wind speed are key drivers of regional climate patterns. Understanding these relationships helps inform climate adaptation strategies for important infrastructure, including airports, in a warming world."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

The results indicate that total precipitation is the most influential predictor, showing a strong inverse relationship with mean temperature. Wind speed and gust speed demonstrate non-linear effects, with moderate increases associated with temperature rises and higher values contributing to cooling. Pressure consistently exhibits a negative association with temperature, with amplified effects at extreme levels. The model explains 61% of the variance in mean temperature and performs well in validation, showing minimal overfitting and low prediction error. These results highlight the role of atmospheric variables in regional temperature changes and the effectiveness of the polynomial regression approach in capturing complex relationships.

@@

In this project, we focus on weather data from YVR Airport, extracting only the datasets containing measurements taken at this specific location from the database. In both datasets, each row corresponds to a single averaged observation for a specific month and year. Each entry includes climate information such as temperature and wind speed, with their respective units recorded alongside the values. Additionally, a unique station ID and geographic coordinates (x, y) are included at the beginning of each entry for reference. The column headers of the raw historical weather dataset is displayed in @tbl-raw-climate. The column headers of the AHCCD dataset is displayed in @tbl-raw-ahccd.

@@

### Model Validation {#sec-validation}

@@

#### Assumptions Fit the Data

@@

#### Validation on Test Data

@@

### Model Justification

>>>>

---
title: "Exploring Nonlinear Atmospheric Influences on Temperature Using Polynomial Regression at Vancouver International Airport"
subtitle: "Pressure and Wind as Key Drivers of Temperature Variability, with Limited Impact from Precipitation and Gust Speed"
author: 
  - Lexun Yu
thanks: "Code and data are available at: [https://github.com/yulexun/ClimateChangeYVR](https://github.com/yulexun/ClimateChangeYVR)."
date: today
date-format: long
abstract: "This study investigates how atmospheric factors influence temperature at Vancouver International Airport using a polynomial regression model. The analysis shows that total precipitation has the strongest effect, significantly reducing mean temperature, while wind speed and atmospheric pressure exhibit non-linear and consistent negative impacts, respectively. By explaining 61% of the variability in temperature, the model demonstrates that precipitation, pressure, gust speed and wind speed are key drivers of regional climate patterns. Understanding these relationships helps inform climate adaptation strategies for important infrastructure, including airports, in a warming world."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

The results indicate that total precipitation is the most influential predictor, showing a strong inverse relationship with mean temperature. Wind speed and gust speed demonstrate non-linear effects, with moderate increases associated with temperature rises and higher values contributing to cooling. Pressure consistently exhibits a negative association with temperature, with amplified effects at extreme levels. The model explains 61% of the variance in mean temperature and performs well in validation, showing minimal overfitting and low prediction error.

These results highlight the role of atmospheric variables in regional temperature changes and the effectiveness of the polynomial regression approach in capturing complex relationships. Our findings contribute to a deeper understanding of regional climate changes, which is important for decision-making and resilience planning in an important nodes in global transportation and logistics, where temperature fluctuations directly influence operational efficiency, safety, and infrastructure planning.

@@

In this project, we focus on weather data from YVR Airport, extracting only the datasets containing measurements taken at this specific location from the database obtained from @canadian_centre_for_climate_services_adjusted_2022 and @meteorological_service_of_canada_past_2023. In both datasets, each row corresponds to a single averaged observation for a specific month and year. Each entry includes climate information such as temperature and wind speed, with their respective units recorded alongside the values. Additionally, a unique station ID and geographic coordinates (x, y) are included at the beginning of each entry for reference. The column headers of the raw historical weather dataset is displayed in @tbl-raw-climate. The column headers of the AHCCD dataset is displayed in @tbl-raw-ahccd.

@@

## Model Validation {#sec-validation}

@@

### Assumptions Fit the Data

@@

### Validation on Test Data

@@

## Model Justification
